Re-examining cross-cultural similarity judgments using lexical co-occurrence

### 2023-01-26 code/4_paper/manuscript/manuscript.tex
changed analysis to exploratory version

\title{Re-examining cross-cultural similarity judgements using lexical
co-occurrence}

@@

\author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA \AND {\large \bf Sharon J.~Derry (SDJ@Macc.Wisc.Edu)} \\ Department of Educational Psychology, 1025 W. Johnson Street \\ Madison, WI 53706 USA}

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or ``chicken''? Speakers of
different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgements toward taxonomic relations, like cow-chicken. This difference
in notions of similarity is the consensus interpretation for
cross-cultural variation in this paradigm. We consider, and provide
evidence for, an alternative possibility, by which notions of similarity
are similar across contexts, but the statistics of the environment vary.
On this account, similarity judgements are guided by co-occurrence in
experience, and observing or hearing about cows and grass or cows and
chickens more often could induce preferences for these groupings, and
account in part for apparent differences in notions of similarity across
contexts.

@@

Systems like fastText and word2vec have been demonstrated to be good
predictors for similarity judgments. Liu, Feng, Wu, Chan, \& Fulton
(2019) showed that fastText performed well as the data for a system
matching responses in a knowledge base by how similar they are to a
customer's query. A study by Jatnika, Bijaksana, \& Suryani (2019)
tested a word2vec model trained on the English wikipedia corpus against
353 word pairs labeled with similarity values based on human judgment.
The word2vec model correlates moderately with human judgment. fastText
has also been shown to be sensitive towards cultural effects on word
meanings: Thompson \& Lupyan (2020) applied fastText to
language-specific Wikipedia corpus (among others) to generate a
``semantic neighborhood'' of 1010 meanings in different languages. The
study showed that languages that are spoken by more similar cultures,
more geographically proximate and/or historically related had more
semantic neighborhoods that aligned. This is a proof of concept for our
use of fastText in different languages with varying levels of
relatedness.

@@

We excluded 26 US participants, 123 Vietnam participants, and 34 China
participants who did not answer all attention checks correctly. We
followed 4 exclusion criteria that aim to retain only participants who
are influenced by one culture: (1) non-native speakers of English and
Vietnamese, respectively, (2) fluent in at least one of the other two
study languages (Vietnamese for US participants, English for Vietnamese
participants and Chinese participants), (3) have lived outside of the
test country (US, Vietnam, or China) for more than two years, and (4)
have significant international experience (more than 6 international
experiences of 2 days or longer.) We did not use a particular criterion
for a language if it would exclude 25\% or more of any one sample. In
this round of exclusion, we excluded 65 US, 19 Vietnam participants, and
34 China participants. After these exclusions, the US sample included
109 participants (27M, 78F, 3 non-binary, 1 other), with mean age = 22.4
(SD = 8.49) and median age = 20. The Vietnam sample included 57
participants (17M, 40F, other), with mean age = 22.21 (SD = 6.12) and
median age = 21. The China sample included 132 participants (51M, 81F,
other), with mean age = 23.27 (SD = 3.75) and median age =
23.\footnote{A table summarizing number of participants lost at each
  round of exclusions is included in the Appendix.}

Notably, we lost a majority of our Vietnam participants (more than 60\%)
in the attention check exclusion. One reason we suspect why this might
have happened is because our Vietnam participants are less familiar with
research surveys and attention check questions, and thus might have
thought too much about the attention check questions.\footnote{We
  carried out an exploratory analysis where we used a less stringent
  attention check exclusion (covered in the final part of the Results \&
  Analysis section). To preview our findings for this analysis, we did
  not find any difference in significant results when compared with the
  main analysis.}

@@

The group means of proportion of thematic response in mainland China is
the highest (M = 0.65, SD = 0.12), followed by the groups means in
Vietnam (M = 0.6, SD = 0.13), which is slightly higher than that of the
US (M = 0.56, SD = 0.17) (Figure 1).

@@

Overall, there is a significant effect of country on proportion of
thematic responses (\(\chi^2\)(2) = 15.22, p \textless{} .001). However,
this effect is driven by the difference between US and China responding
(\(\beta\) = -0.53, p \textless{} .001). There is no statistical
difference between the Vietnam and China responding (\(\beta\) = -0.25,
p = 0.137), and the US and Vietnam responding (\(\beta\) = 0.28, p =
0.142).

@@

We found that all corpora are significant predictors of all cultural
context responding, with p\textless0.05 and \(\beta\) from -9.05 to
-2.44. (For a full report, see Appendix.)

@@

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -7.14,
\(\chi^2\)(1) = 16.78, p \textless{} .001. VI corpus: \(\beta\) = -2.28,
\(\chi^2\)(1) = 3.74, p = 0.053. ZH corpus: \(\beta\) = -3.53,
\(\chi^2\)(1) = 3.99, p = 0.046.

For Vietnam responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -4.25,
\(\chi^2\)(1) = 3.87, p = 0.049. VI corpus: \(\beta\) = -2.98,
\(\chi^2\)(1) = 4.17, p = 0.041. ZH corpus: \(\beta\) = -4.86,
\(\chi^2\)(1) = 4.94, p = 0.026.

For China responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -3.52,
\(\chi^2\)(1) = 7.85, p = 0.005. VI corpus: \(\beta\) = -1.62,
\(\chi^2\)(1) = 3.57, p = 0.059. ZH corpus: \(\beta\) = -5.35,
\(\chi^2\)(1) = 17.19, p \textless{} .001.

@@

However, we found that language specificity alone does not explain our
results. We used an ANOVA to compare the model with only the
corresponding corpus, and the model with all 3 corpora for each cultural
context. We found that in all three cases (US, China, Vietnam), adding
the other two corpora produces a significantly better fit than the
identical model without the additional corpora, and only the
corresponding corpus included as a predictor (US response: \(\chi^2\)(2)
= 8.42, p = 0.015; Vietnam response: \(\chi^2\)(2) = 15.21, p
\textless{} .001; China response: \(\chi^2\)(2) = 10.83, p NA).

@@

Our analysis shows that lexical statistics of each corpus is a
significant predictor for similarity judgement in the corresponding
cultural context, even without the addition of the other two corpora.
Assuming that lexical statistics is a good proxy for environmental
statistics, we take the above result as evidence that cross-cultural
differences in input to the process of similarity judgement at least
partially drives cross-cultural similarity judgement differences.
However, it is still an open question whether the nature of similarity
judgement (for e.g., the notion of similarity) itself differs across
cultures and contributes to cross-cultural differences in similarity
judgement responding. In an alternative hypothesis, the cross-cultural
differences in similarity judgement observed is driven by not only the
input, but also an interaction between cultural-specific nature of
similarity judgement and input. To test this hypothesis, we
operationalize cultural-specific nature of similarity judgement as
country context, and input to similarity judgement as the corresponding
corpus statistics. We then compare a model that includes country,
corresponding corpus statistics and their interaction as fixed effects
(R syntax: responses \textasciitilde{} corresponding\_corpus * country +
(1\textbar triad) + (1\textbar subject)) to a model with only the
corresponding corpus statistics as the fixed effect (R syntax: responses
\textasciitilde{} corresponding\_corpus + (1\textbar triad) +
(1\textbar subject))

In the model containing only the corresponding corpus statistics, the
corpus statistics is a significant predictor (\(\beta\) = -1.75,
\(\chi^2\)(1) = 48.9, p \textless{} .001). When adding country and
interaction between country and corpus, both the corpus statistics
(\(\beta\) = -1.51, \(\chi^2\)(1) = 13.96, p \textless{} .001) and
interaction between country and corpus (\(\beta\) = NA, \(\chi^2\)(2) =
9.1, p = 0.011) are significant predictors. This suggests that there are
cross-cultural differences in both input and nature of similarity
judgement that drive differences in similarity judgement responding.

@@

Previous work using the triad task such as Ji et al. (2004) has focused
on contrasting taxonomic-thematic match preferences, driven by a theory
of cross-cultural differences in notions of similarity. Meanwhile, our
approach has shown that notions of similarity only has an effect on
cross-cultural differences in similarity judgement when modulated by
input, and that input as proxied by lexical statistics of the
corresponding corpus by itself is a significant predictor for similarity
judgement responding. Our approach using lexical statistics do not
assume a taxonomic/thematic structure in the triads. Therefore, it
should be possible to predict similarity judgements to triads that do
not follow a taxonomic/thematic structure, such as our filler triads. If
our approach is tractable, we should be able to repeat the above
analyses to the filler triads, and find that lexical statistics of each
corpus is a significant predictor for similarity judgements in filler
triads of the corresponding cultural context.

@@

\caption[Proportion of thematic responses by country]{Proportion of thematic responses by country.}\label{fig:unnamed-chunk-4}
\end{figure}
\end{CodeChunk}

@@

Using the same mixed-effects logistic regression structure as Analysis
2, we predict responses (1=word1 or 0=word2) with corpus prediction
(proportion of cosine distance) as a fixed effect and participant and
triad as random effects. We found that in all cultural contexts, the
correpsonding corpora is a significant predictor of responding, with
p\textless0.05 and \(\beta\) from -9.59 to -3.05. (For a full report,
see Appendix.)

@@

In this paper, we consider whether statistics of the language
environment can account for cross-cultural differences in a classic
similarity judgment paradigm, as an alternative to the view that members
of different cultures vary in their conception of similarity.

@@

We did find some signatures of language specificity in our analysis,
such as the large positive correlation between similarity judgments of
each country and the respective corpus statistics, and how each corpus
statistics are good predictors for corresponding country's similarity
judgments. However, this is potentially due to the high correlation
between corpus statistics of English, Vietnamese and Mandarin. We find
even stronger evidence for consistency across the three groups, with
substantive overlapping predictions across the corpus models, highly
similar responding across the experiments, and a correspondingly high
fit in cross-language comparisons between models and data.

Our findings raise additional questions for future work: if not
differences in taxonomic vs.~thematic responding, then what differences
drive the relativity effects previous studies have observed? To what
extent are the relativity effects driven by language, and to what extent
by culture? Ji et al. (2004) established that culture-aligned
differences in this paradigm exist, even when the test language is held
constant, concluding that ``it is culture (independent of the testing
language) that led to different grouping styles'' in their study. Our
data provide a cautionary note to this conclusion, suggesting that
semantic representations in bilinguals (see Francis (2005) for a review)
may have the potential to provide an offline account for cross-context
differences in similarity judgments, independent of test language.
However, there are still many open questions for this account. How do
semantic associations guide categorization? Can they explain
taxonomic-thematic differences of the type reported by Ji et al. (2004)
and others? Can we provide a more specific computational account than
the simple frequency model tested here?

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
Sample & Intra-rater ICC & Inter-rater ICC \\ 
  \hline
US & 0.32 & 0.97 \\ 
  China & 0.20 & 0.96 \\ 
  Vietnam & 0.32 & 0.95 \\ 
   \hline
\end{tabular}
\caption{ICCs for reliability within participants (intra-rater) and stimuli (inter-rater) for all cultural contexts.} 
\end{table}

\hypertarget{supplemental-for-whether-each-corpus-predicts-similarity-judgement-in-each-cultural-context}{%
\subsection{2. Supplemental for whether each corpus predicts similarity
judgement in each cultural
context}\label{supplemental-for-whether-each-corpus-predicts-similarity-judgement-in-each-cultural-context}}

@@

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are significant predictors of US responding (EN-US model:
\(\beta\) = -9.05, \(\chi^2\)(1) = 31.28, p \textless{} .001; VI-US
model: \(\beta\) = -3.13, \(\chi^2\)(1) = 5.1, p = 0.024; ZH-US model:
\(\beta\) = -7.42, \(\chi^2\)(1) = 18.48, p \textless{} .001).

For Vietnam (VN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of VN responding (EN-VN model:
\(\beta\) = -6.86, \(\chi^2\)(1) = 11.71, p \textless{} .001; VI-VN
model: \(\beta\) = -3.83, \(\chi^2\)(1) = 6.02, p = 0.014; ZH-VN model:
\(\beta\) = -7.49, \(\chi^2\)(1) = 13.84, p \textless{} .001).

For China (CN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of CN responding (EN-CN model:
\(\beta\) = -6.18, \(\chi^2\)(1) = 25.39, p \textless{} .001; VI-CN
model: \(\beta\) = -2.44, \(\chi^2\)(1) = 5.75, p = 0.017; ZH-CN model:
\(\beta\) = -7.36, \(\chi^2\)(1) = 37.72, p \textless{} .001).

@@

English (EN) is a significant predictors of US responding (\(\beta\) =
-7.93, \(\chi^2\)(1) = 35.87, p \textless{} .001).

Vietnamese (VI) is a significant predictors of Vietnam responding
(\(\beta\) = -3.05, \(\chi^2\)(1) = 5.88, p = 0.015).

@@

\hypertarget{demographic-exclusion-statistics}{%
\subsection{3. Demographic Exclusion
Statistics}\label{demographic-exclusion-statistics}}

@@

>>>>

\title{Re-examining cross-cultural similarity judgments using lexical
co-occurrence}

@@

\author{{\large \bf Khuyen N. Le (knl005@ucsd.edu)}$^{1\star}$, {\large \bf Alexandra Carstensen (abc@ucsd.edu)}$^{1\star}$, \\ {\large \bf Shan Gao (shangaocog@gmail.com)$^2$} {\large \bf Michael C. Frank (mcfrank@stanford.edu)$^3$},  \AND $^1$Department of Psychology, UC San Diego, $^2$Department of Psychology, University of Chicago, \\ $^3$Department of Psychology, Stanford University }

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or ``chicken''? Speakers of
different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow-chicken. This difference
in notions of similarity is the consensus interpretation for
cross-cultural variation in this paradigm. We consider, and provide
evidence for, an alternative possibility, by which notions of similarity
are similar across contexts, but the statistics of the environment vary.
On this account, similarity judgments are guided by co-occurrence in
experience, and observing or hearing about cows and grass or cows and
chickens more often could induce preferences for these groupings, and
account in part for apparent differences in notions of similarity across
contexts.

@@

Word embeddings like fastText have been demonstrated to be good
predictors for similarity judgments (Liu, Feng, Wu, Chan, \& Fulton
(2019), Jatnika, Bijaksana, \& Suryani (2019)). fastText has also been
shown to be sensitive towards cultural effects on word meanings:
Thompson \& Lupyan (2020) showed that the distribution of semantic
meaning clusters generated by fastText trained on language-specific
corpora correlates to the cultural, historical and geographical
distances of such languages languages. This is a proof of concept for
our use of fastText in different languages with varying levels of
relatedness.

@@

We excluded 8 US participants, 62 Vietnam participants, and 16 China
participants who did not answer 2 or more attention checks correctly. We
followed 4 exclusion criteria that aim to retain only participants who
are influenced by one culture: (1) non-native speakers of English and
Vietnamese, respectively, (2) fluent in at least one of the other two
study languages (Vietnamese for US participants, English for Vietnamese
participants and Chinese participants), (3) have lived outside of the
test country (US, Vietnam, or China) for more than two years, and (4)
have significant international experience (more than 6 international
experiences of 2 days or longer.) We did not use a particular criterion
for a language if it would exclude 25\% or more of any one sample. In
this round of exclusion, we excluded 73 US, 27 Vietnam participants, and
35 China participants. After these exclusions, the US sample included
119 participants (30M, 84F, 3 non-binary, 2 other), with mean age = 22.2
(SD = 8.15) and median age = 20. The Vietnam sample included 110
participants (34M, 71F, 5 other), with mean age = 22.21 (SD = 5.81) and
median age = 21. The China sample included 149 participants (61M, 87F, 1
other), with mean age = 23.1 (SD = 3.65) and median age = 23.\footnote{A
  table summarizing number of participants lost at each round of
  exclusions is included in the Appendix.}

Notably, we lost a majority of our Vietnam participants (more than 60\%)
in the attention check exclusion. One reason we suspect why this might
have happened is because our Vietnam participants are less familiar with
research surveys and attention check questions, and thus might have
thought too much about the attention check questions.\footnote{We
  carried out an exploratory analysis where we used a less stringent
  attention check exclusion, and did not find any difference in
  significant results when compared with the main analysis (see
  Appendix).}

@@

The group means of proportion of thematic response in mainland China is
the highest (M = 0.65, SD = 0.11), followed by the groups means in
Vietnam (M = 0.6, SD = 0.11), which is slightly higher than that of the
US (M = 0.56, SD = 0.17) (Figure 1).

@@

Overall, there is a significant effect of country on proportion of
thematic responses (\(\chi^2\)(2) = 15.37, p \textless{} .001). However,
this effect is driven by the difference between US and China responding
(\(\beta\) = -0.48, p \textless{} .001). There is no statistical
difference between the Vietnam and China responding (\(\beta\) = -0.22,
p = 0.09), and the US and Vietnam responding (\(\beta\) = 0.26, p =
0.086).

@@

We found that all corpora are significant predictors of all cultural
context responding, with p \textless{} 0.05 and \(\beta\) from -8.69 to
-2.4. (For a full report, see Appendix.)

@@

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -6.96,
\(\chi^2\)(1) = 16.57, p \textless{} .001. VI corpus: \(\beta\) = -2.22,
\(\chi^2\)(1) = 3.73, p = 0.054. ZH corpus: \(\beta\) = -3.18,
\(\chi^2\)(1) = 3.37, p = 0.066.

For Vietnam responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -2.98,
\(\chi^2\)(1) = 2.58, p = 0.108. VI corpus: \(\beta\) = -2.75,
\(\chi^2\)(1) = 4.84, p = 0.028. ZH corpus: \(\beta\) = -4.39,
\(\chi^2\)(1) = 5.49, p = 0.019.

For China responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = -3.32,
\(\chi^2\)(1) = 7.18, p = 0.007. VI corpus: \(\beta\) = -1.59,
\(\chi^2\)(1) = 3.63, p = 0.057. ZH corpus: \(\beta\) = -5.31,
\(\chi^2\)(1) = 17.69, p \textless{} .001.

@@

However, we found that language specificity alone does not explain our
results. We used an ANOVA to compare the model with only the
corresponding corpus, and the model with all 3 corpora for each cultural
context. We found that in all three cases (US, China, Vietnam), adding
the other two corpora produces a significantly better fit than the
identical model without the additional corpora, and only the
corresponding corpus included as a predictor (US response: \(\chi^2\)(2)
= 7.93, p = 0.019; Vietnam response: \(\chi^2\)(2) = 13.72, p = 0.001;
China response: \(\chi^2\)(2) = 10.56, p = 0.005).

@@

Our analysis shows that lexical statistics of each corpus is a
significant predictor for similarity judgment in the corresponding
cultural context, even without the addition of the other two corpora.
Assuming that lexical statistics is a good proxy for environmental
statistics, we take the above result as evidence that cross-cultural
differences in input to the process of similarity judgment at least
partially drives cross-cultural similarity judgment differences.
However, it is still an open question whether the nature of similarity
judgment (for e.g., the notion of similarity) itself differs across
cultures and contributes to cross-cultural differences in similarity
judgment responding. In an alternative hypothesis, the cross-cultural
differences in similarity judgment observed is driven by not only the
input, but also an interaction between cultural-specific nature of
similarity judgment and input. To test this hypothesis, we
operationalize cultural-specific nature of similarity judgment as
country context, and input to similarity judgment as the corresponding
corpus statistics. We then compare a model that includes country,
corresponding corpus statistics and their interaction as fixed effects
(R syntax: responses \textasciitilde{} corresponding\_corpus * country +
(1\textbar triad) + (1\textbar subject)) to a model with only the
corresponding corpus statistics as the fixed effect (R syntax: responses
\textasciitilde{} corresponding\_corpus + (1\textbar triad) +
(1\textbar subject))

In the model containing only the corresponding corpus statistics, the
corpus statistics is a significant predictor (\(\beta\) = -1.79,
\(\chi^2\)(1) = 75.58, p \textless{} .001). When adding country and
interaction between country and corpus, both the corpus statistics
(\(\beta\) = -1.72, \(\chi^2\)(1) = 23.34, p \textless{} .001) and
interaction between country and corpus (F(2, \ensuremath{\infty{}}) =
6.66, \(\chi^2\)(2) = 13.33, p = 0.001) are significant predictors. This
suggests that there are cross-cultural differences in both input and
nature of similarity judgment that drive differences in similarity
judgment responding. We found no effects of country (F(2,
\ensuremath{\infty{}}) = 13.18, \(\chi^2\)(2) = 7.39, p = 0.025).

@@

Previous work using the triad task such as Ji et al. (2004) has focused
on contrasting taxonomic-thematic match preferences, driven by a theory
of cross-cultural differences in notions of similarity. Meanwhile, our
approach has shown that notions of similarity only has an effect on
cross-cultural differences in similarity judgment when modulated by
input, and that input as proxied by lexical statistics of the
corresponding corpus by itself is a significant predictor for similarity
judgment responding. Our approach using lexical statistics do not assume
a taxonomic/thematic structure in the triads. Therefore, it should be
possible to predict similarity judgments to triads that do not follow a
taxonomic/thematic structure, such as our filler triads. If our approach
is tractable, we should be able to repeat the above analyses to the
filler triads, and find that lexical statistics of each corpus is a
significant predictor for similarity judgments in filler triads of the
corresponding cultural context.

@@

\caption[Proportion of Word1 responses by country]{Proportion of Word1 responses by country.}\label{fig:unnamed-chunk-4}
\end{figure}
\end{CodeChunk}

@@

Using the same mixed-effects logistic regression structure as Analysis
2, we predict responses (1=word1 or 0=word2) with corpus prediction
(proportion of cosine distance) as a fixed effect and participant and
triad as random effects. We found that in all cultural contexts, the
correpsonding corpora is a significant predictor of responding, with
p\textless0.05 and \(\beta\) from -9.59 to -3.14. (For a full report,
see Appendix.)

@@

In this paper, we consider whether statistics of the language
environment can account for cross-cultural differences in a classic
similarity judgment paradigm, as an alternative to the view that members
of different cultures vary in their conception of similarity. We found
that each corpus statistics is a good predictor for the corresponding
country's similarity judgments. When corresponding corpus statistics,
country and corpus-country interactions were included as fixed effects
to predict similarity judgments, we found no effects of country, but
both corpus statistics and corpus-country interaction are significant
predictors. Our results provide evidence that cross-cultural differences
in similarity judgment might not be driven by cross-cultural differences
in conception of similarity alone, but rather by statistics to
similarity judgment and its interaction with similarity conception. We
also showed that corpus statistics predict similarity judgments for
triads that do not have taxonomic/thematic structure.

@@

We did find some signatures of language specificity in our analysis:
each corpus statistics is a good predictor for the corresponding
country's similarity judgments, even when other corpus statistics are
included. However, we also find evidence for consistency across the
three cultural groups: each corpus statistics is also a good predictor
for the other two non-corresponding cultural contexts when included as
the only predictor, and both English and Mandarin Chinese statistics are
good predictors for their non-corresponding cultural contexts when all
three corpora are included.

When all cultural contexts are taken together, we found that both
corresponding corpus statistics and an interaction between corpus and
country are significant predictors for similarity judgments. This result
suggests that in addition to input to similarity judgment,
country-specific categorical differences in similarity judgment (for
e.g., how each culture conceive of similarity) also predict
cross-cultural differences in similarity judgment behavior while
modulated by input.

Even in triads with no taxonomic / thematic structure, each corpus
statistics is still a good predictor of the corresponding country's
similarity judgments. The ability of corpus statistics to predict
similarity judgments in non-structured triads suggests that
cross-cultural differences in similarity judgments should be construed
in a more general and graded manner, as opposed to binary
taxonomic/thematic match-preferring.

Our findings raise additional questions for future work: To what extent
are the relativity effects driven by language, and to what extent by
culture? Ji et al. (2004) established that culture-aligned differences
in this paradigm exist, even when the test language is held constant,
concluding that ``it is culture (independent of the testing language)
that led to different grouping styles'' in their study. Our data provide
a cautionary note to this conclusion, suggesting that semantic
representations in bilinguals (see Francis (2005) for a review) may have
the potential to provide an offline account for cross-context
differences in similarity judgments, independent of test language.
However, there are still many open questions for this account. How do
semantic associations guide categorization? Can they explain
taxonomic-thematic differences of the type reported by Ji et al. (2004)
and others? Can we provide a more specific computational account than
the simple frequency model tested here?

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
Sample & Intra-rater ICC & Inter-rater ICC \\ 
  \hline
US & 0.30 & 0.97 \\ 
  China & 0.19 & 0.96 \\ 
  Vietnam & 0.27 & 0.97 \\ 
   \hline
\end{tabular}
\caption{ICCs for reliability within participants (intra-rater) and stimuli (inter-rater) for all cultural contexts.} 
\end{table}

An intraclass correlation coefficients (ICCs) analysis with a two-way,
non-interactive model (following the standard taxonomy for ICCs from
McGraw \& Wong (1996)) showed that, per cultural context, participants'
judgments were variable across triads, while inter-rater agreement about
specific triads was quite high (Table 1). This result suggests that each
triad prompts taxonomic / thematic matching reliably within culture.

\hypertarget{supplemental-for-whether-each-corpus-predicts-similarity-judgment-in-each-cultural-context}{%
\subsection{2. Supplemental for whether each corpus predicts similarity
judgment in each cultural
context}\label{supplemental-for-whether-each-corpus-predicts-similarity-judgment-in-each-cultural-context}}

@@

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are significant predictors of US responding (EN-US model:
\(\beta\) = -8.69, \(\chi^2\)(1) = 30.98, p \textless{} .001; VI-US
model: \(\beta\) = -3.01, \(\chi^2\)(1) = 5.24, p = 0.022; ZH-US model:
\(\beta\) = -6.97, \(\chi^2\)(1) = 17.62, p \textless{} .001).

For Vietnam (VN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of VN responding (EN-VN model:
\(\beta\) = -5.33, \(\chi^2\)(1) = 9.4, p = 0.002; VI-VN model:
\(\beta\) = -3.46, \(\chi^2\)(1) = 6.8, p = 0.009; ZH-VN model:
\(\beta\) = -6.37, \(\chi^2\)(1) = 13.84, p \textless{} .001).

For China (CN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of CN responding (EN-CN model:
\(\beta\) = -5.97, \(\chi^2\)(1) = 25.49, p \textless{} .001; VI-CN
model: \(\beta\) = -2.4, \(\chi^2\)(1) = 5.91, p = 0.015; ZH-CN model:
\(\beta\) = -7.22, \(\chi^2\)(1) = 39.24, p \textless{} .001).

@@

English (EN) is a significant predictors of US responding (\(\beta\) =
-8.35, \(\chi^2\)(1) = 36.1, p \textless{} .001).

Vietnamese (VI) is a significant predictors of Vietnam responding
(\(\beta\) = -3.14, \(\chi^2\)(1) = 4.65, p = 0.031).

@@

\hypertarget{exclusion-statistics}{%
\subsection{3. Exclusion Statistics}\label{exclusion-statistics}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-McGraw1996}{}}%
McGraw, K. O., \& Wong, S. P. (1996). Forming inferences about some
intraclass correlation coefficients. \emph{Psychological Methods},
\emph{1}(1), 30.
